Theories of deliberated choice

### 2024-01-22 simple.tex
Rm wrong conjecture about (not) HAI

    For \cref{it:HAI} $\implies$ \cref{it:condHAI},
    pick an indicative claim $C$ and consider $i \in I$ and $\theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub$ such that $\pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
    Define $H = \theta^{-1}(i, P)$.
  \end{proof}
  \begin{corollary}
    \label{th:suffIndicDec}
    Assume that for some $p \in P$, we have $\exists i \in I \suchthat P_i = p$ and $\forall i_2 \suchthat P_{i_2} \neq p$ we have that $\exists A \in \finite(\allargs) \suchthat \forall i \in I \suchthat P_i \neq p, \exists a \in A \suchthat {\ileadstost}(a) \in P \setminus \set{p}$, then there exists a claim that holds and is anonymous and indicative.
  \end{corollary}
  \begin{proof}
    We know that $i_1$ has some decisive argument $a_1$ in favor of $p$.
    Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
    Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist_i = (\hist_{i - 1}, a_1, i)$.
    Define $H = \set{\hist_i \suchthat i \in \intvl{2, \card{A} + 1}}$.
    Apply \cref{th:suffIndic}.
  \end{proof}
\begin{example}
  Consider $I_p$ such that $\forall i \in I_p$: $i$ constantly says $p$.
  Consider $J = I \setminus I_p$ whose DJ is $p' \neq p$, and such that each $j \in J$ has a single decisive argument $\ar_j$ for $p'$, and when $\ar_j$ is not in $\hist$, $\hist \ileadsto[j] p$.
  Then, the condition is not satisfied.
\end{example}
\begin{conjecture}
  \label{th:caractNotIndicWrong}
  The following two conditions on $\allleadsto$ are equivalent:
  \begin{enumerate}[label=T\ref{th:caractNotIndic}-{\arabic*}]
    \item \label{it:NotHAIWrong} there exists no claim that holds and is anonymous and indicative;
    \item \label{it:condNotHAIWrong} either all individuals have an empty deliberated preference, or all paths are such that at least two individuals with different deliberated judgements behave identically on the whole path (formally: either $\pi(I) = \emptyset$ or $\forall \hist \in \allhists, \exists i_1, i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$);
    \item \label{it:condNotHAIInfWrong} [To be adjusted for the zero case] all paths are such that infinitely many individuals with different deliberated judgements behave identically on the whole path (formally: $\forall \hist \in \allhists, \exists I_1 \wsub I, i_2 \in I \suchthat \card{I_1} = \infty \land \forall i_1 \in I_1: \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$).
  \end{enumerate}

  This is WRONG because could have $i$ that says $\phi$ on $H$ and has $\phi$ as DJ, and $i$ can be isolated (nobody with DJ different than $\phi$ says $\phi$ on $H$) but still some $i_1, i_2$ with different DJ behave identically on $H$, say they both say $¬\phi$ on $H$.
\end{conjecture}
\begin{proof}
  Let us show that NOT \ref{it:condNotHAIWrong} implies \ref{it:condHAI}.
  Consider: at least one individual $i$ has a non-empty DP, and some path is such that
  $\forall i_1, i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2): \exists \histp \wsub \hist \suchthat {\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $H = \set{\histp \wsub \hist}$ and $i_1 = i$.

  Conversely, \ref{it:condHAI} implies NOT \ref{it:condNotHAIWrong}.
  Consider: for some $H \in \finite(\incr(\allhists))$ and $i \in I \suchthat P_{i} \in P$, we have $\forall i_2 \suchthat P_{i_2} \neq P_i$ that $\exists \histp \in H \suchthat {\ileadsto}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $\hist$ as the maximal path (in cardinal) in $H$.
  Consider any $i_1, i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2)$.
  Remains to show that for some $\histp \wsub \hist$: ${\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Assume, on the contrary, that for all $\histp \wsub \hist$: ${\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$.
\end{proof}
\begin{conjecture}
  \label{th:caractNotIndic}
  The following conditions on $\allleadsto$ are equivalent:
  \begin{enumerate}[label=T\ref{th:caractNotIndic}-{\arabic*}]
    \item \label{it:NotHAI} there exists no claim that holds and is anonymous and indicative;
    \item \label{it:condNotHAI} all paths are such that for all $i_1$ having a DJ, there is an $i_2$ with another deliberated judgement that behaves identically on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I \suchthat \pi(i_1) \neq \emptyset: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$);
    \item \label{it:condNotHAIInf} all paths are such that for all $i_1$, either at least one individual without a DJ, or infinitely many individuals $i_2$ with another deliberated judgement than $i_1$, behave identically to $i_1$ on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I: [\exists i_2 \in I \suchthat \pi(i_2) = \emptyset \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)] \lor [\card(\set{i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}) = \infty]$)
  \end{enumerate}
\end{conjecture}
\begin{proof}
  Let us show that NOT \ref{it:condNotHAI} implies \ref{it:condHAI}.
  Consider: some individual $i$ and some path $\hist$ are such that $\pi(i) \neq \emptyset$ and
  $\forall i_2 \in I \suchthat \pi(i) \neq \pi(i_2): \exists \histp \wsub \hist \suchthat {\ileadsto[i]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $H = \set{\histp \wsub \hist}$ and $i_1 = i$.

>>>>

    For \cref{it:HAI} $\implies$ \cref{it:condHAI},
    pick an indicative claim $C$ and consider $i \in I$ and $\theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub$ such that $\pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
    Define $H = \theta^{-1}(i, P)$.
  \end{proof}
  \begin{corollary}
    \label{th:suffIndicDec}
    Assume that for some $p \in P$, we have $\exists i \in I \suchthat P_i = p$ and $\forall i_2 \suchthat P_{i_2} \neq p$ we have that $\exists A \in \finite(\allargs) \suchthat \forall i \in I \suchthat P_i \neq p, \exists a \in A \suchthat {\ileadstost}(a) \in P \setminus \set{p}$, then there exists a claim that holds and is anonymous and indicative.
  \end{corollary}
  \begin{proof}
    We know that $i_1$ has some decisive argument $a_1$ in favor of $p$.
    Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
    Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist_i = (\hist_{i - 1}, a_1, i)$.
    Define $H = \set{\hist_i \suchthat i \in \intvl{2, \card{A} + 1}}$.
    Apply \cref{th:suffIndic}.
  \end{proof}
\begin{example}
  Consider $I_p$ such that $\forall i \in I_p$: $i$ constantly says $p$.
  Consider $J = I \setminus I_p$ whose DJ is $p' \neq p$, and such that each $j \in J$ has a single decisive argument $\ar_j$ for $p'$, and when $\ar_j$ is not in $\hist$, $\hist \ileadsto[j] p$.
  Then, the condition is not satisfied.
\end{example}
\begin{conjecture}
  \label{th:caractNotIndic}
  The following conditions on $\allleadsto$ are equivalent:
  \begin{enumerate}[label=T\ref{th:caractNotIndic}-{\arabic*}]
    \item \label{it:NotHAI} there exists no claim that holds and is anonymous and indicative;
    \item \label{it:condNotHAI} all paths are such that for all $i_1$ having a DJ, there is an $i_2$ with another deliberated judgement that behaves identically on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I \suchthat \pi(i_1) \neq \emptyset: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$);
    \item \label{it:condNotHAIInf} all paths are such that for all $i_1$, either at least one individual without a DJ, or infinitely many individuals $i_2$ with another deliberated judgement than $i_1$, behave identically to $i_1$ on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I: [\exists i_2 \in I \suchthat \pi(i_2) = \emptyset \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)] \lor [\card(\set{i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}) = \infty]$)
  \end{enumerate}
\end{conjecture}
\begin{proof}
  Let us show that NOT \ref{it:condNotHAI} implies \ref{it:condHAI}.
  Consider: some individual $i$ and some path $\hist$ are such that $\pi(i) \neq \emptyset$ and
  $\forall i_2 \in I \suchthat \pi(i) \neq \pi(i_2): \exists \histp \wsub \hist \suchthat {\ileadsto[i]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $H = \set{\histp \wsub \hist}$ and $i_1 = i$.
